---
title: "Regression Models Project"
author: "Ruel Laranjo | Junior Data Scientist "
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary

This project analyzes whether transmission type (manual vs automatic) has a significant impact on fuel efficiency (miles per gallon, MPG). Using the mtcars dataset, a linear regression model was applied to quantify the relationship between transmission type and MPG.

The results suggest that manual transmission vehicles tend to have higher fuel efficiency compared to automatic vehicles. However, this relationship is influenced by other factors such as vehicle weight and horsepower.

This analysis demonstrates how regression modeling can be used to identify key variables affecting performance and support data-driven decision-making.

# Interpretation of Results

The regression model indicates that transmission type has a measurable effect on fuel efficiency. Manual cars generally show higher MPG values compared to automatic cars.

However, when additional variables such as weight and horsepower are included, the difference between transmission types becomes less pronounced. This suggests that vehicle characteristics play a significant role in determining fuel efficiency.

Overall, the model highlights the importance of considering multiple variables when analyzing real-world data.

# Data Exploration

```{r}
data(mtcars)
head(mtcars)
```

# Regression Analysis

```{r}
model1 <- lm(mpg ~ am, data = mtcars)
summary(model1)
```

The coefficient for am represents the average difference in MPG between manual and automatic cars. A positive coefficient indicates that manual transmission cars have higher fuel efficiency compared to automatic cars.

```{r}
model2 <- lm(mpg ~ am + wt + hp, data = mtcars)
summary(model2)
```

Model 2 adjusts for weight (wt) and horsepower (hp). Comparing the adjusted R-squared values suggests that Model 2 explains more variability in MPG and provides a better fit than Model 1.

```{r}
par(mfrow=c(2,2))
plot(model2)
```

The residual plots do not show strong systematic patterns, suggesting that the assumptions of linear regression are reasonably satisfied.


```{r}
confint(model2)
```

The confidence interval for the am coefficient indicates the range of plausible values for the MPG difference between transmission types. If the interval does not include zero, the effect is statistically significant.

# Conclusion

This analysis shows that transmission type has an impact on fuel efficiency, with manual vehicles generally achieving higher MPG than automatic vehicles.

However, the results also indicate that other factors such as vehicle weight and horsepower significantly influence fuel efficiency. When these variables are included in the model, the effect of transmission type becomes less dominant.

Overall, this project demonstrates the importance of using multiple variables in regression analysis to obtain more accurate and meaningful insights from data.








